\subsubsection{Vice President,
 Communications}
\label{vice-president-communications}
The Vice President, Communications shall:

\begin{enumerate}
 \item
  Liaise with faculty and university wide groups oriented around the public relations of the MES.
 \item
  Be responsible for maintaining the brand of the MES.
 \item
  Chair the Advertising Committee. 
 \item 
  Hold a seat on the Editorial Review Committee 
 \item 
  Work alongside the President, VPs, Directors, and Appointed Positions to ensure members are well informed of the activities, events, and services of the MES.
 \item
  Create and implement a marketing strategy for MES events, in partnership with the Social Media Coordinator(s), Content Coordinator(s), and Website Coordinator(s).
 \item
  Maintain a respectful relationship with the current advertising and printing agencies.
 \item
  Create and implement a method of obtaining information from the VPs, Directors, Appointed Positions, and Program Representatives for event and service promotion on the MES Website and social media platforms.
 \item
  Work with the positions in their portfolio, as well as any other members of the community to ensure that the website, MES social media accounts, and MES publications are effectively used as communication tools, and positively reflect the MES brand.
 \item
  Actively search for more effective ways for the MES to advertise its events, services, and publications, and adapt advertising methods to the preferred means of the membership.
 \item
  Maintain promotional media for the Hatch Building TVs.
 \item
  Work to improve the sustainability of MES promotional campaigns.
 \item
  Work with the President and WW Co-OCs to create the Summer Mail-Out, the WW promotional campaign, and to develop a social media plan for WW.
 \item
  Organize appropriate MES representation to promote the MES and the engineering profession at events, foremost:

  \begin{enumerate}
   \item
    Faculty of Engineering Open Houses
   \item
    MES charity events
  \end{enumerate}
 \item
  Develop new charity initiatives and encourage the development of activities that will actively promote the image of the MES.
 \item
  Coordinate with McMaster University's Gender Equality Committee to promote the December 6th Memorial.
 \item
  Liaise with the Engineering Alumni Office regarding outreach, alumni, or other relevant initiative.
 \item
  Attend meetings initiated by the Engineering Alumni Office and participate in Alumni events as an MES representative.
 \item
  Report to the President.
 \item
  Supervise the following positions:

  \begin{enumerate}
   \item
    Graphic Designer(s)
   \item
    Publications Editor
   \item
    Content Coordinator(s)
   \item
    Handbook Editor(s)
   \item
    Frequency Editor(s)
   \item   
    Frequency Designer(s) 
   \item    
    Frequency Writer(s) 
   \item
    Plumbline Editor(s)
   \item
    Plumbline Designer(s) 
   \item     
    Plumbline Writer(s) 
   \item     
    Social Media Coordinator(s)
   \item
    Photographer/Videographer(s)
  \end{enumerate}

\end{enumerate}