\subsubsection{Vice President, Student
 Life}
\label{vice-president-student-life}
The Vice President, Student Life shall:

\begin{enumerate}
 \item
  Act as the President in all capacities in the latter's absence.
 \item
  Serve as a member of the following committees:

  \begin{enumerate}
   \item
    Elections Committee (see MES Bylaws Section \ref{elections-committee})
   \item
    Awards Selection Committee (see MES Bylaws Section \ref{awards-selection-committee})
   \item
    Sports Committee (see MES Bylaws Section \ref{sports-committee})
   \item
    Hatch Operations Committee (Faculty committee)
  \end{enumerate}
 \item
  Be an ex-officio member of all MES Committees except those to which they have been duly appointed.
 \item
 Plan and organize Faculty Frenzy during Back2School Week with the assistance of Director(s) of Sports. 
 \item
  Coordinate and organize events for Frost Week.
 \item
  Be responsible for approving any interfaculty morale boosting activities on behalf of the MES.
 \item
  Liaise with University Health and Safety and ensure all MES and Program Society social activities abide by its guidelines.
 \item
  Report to the President.
 \item
  Supervise the following positions:

  \begin{enumerate}
   \item
    Director of Events
   \item
    Director(s) of Wellness
   \item
    Director(s) of Sustainability   
   \item
    Director(s) of Sports    
   \item
    Fireball Coordinators 
   \item   
    The Phoenix Formal (previously known as Graduation Formal) Coordinators   
   \item    
    Any Pitch Your Own Position deemed an appropriate fit under Student Life portfolio by President and Vice-President, Student Life 
  \end{enumerate}
 \item
  Develop protocols, with any members at the VPSL's discretion, for the purposes of risk aversion and incident mitigation for all aspects of MES events. To be completed by September.

\end{enumerate}