\section{Structure \& Organization}
\label{structure-organization}

\subsection{The Executive}
\label{the-executive}

The MES Executive shall consist of the following positions. One person shall occupy each position.

\input{Sections/Exec Portfolios/president.tex}
\input{Sections/Exec Portfolios/vice-president-student-life.tex}
\input{Sections/Exec Portfolios/vice-president-external-relations.tex}
\input{Sections/Exec Portfolios/vice-president-academic.tex}
\input{Sections/Exec Portfolios/vice-president-finance.tex}
\input{Sections/Exec Portfolios/vice-president-internal.tex}
\input{Sections/Exec Portfolios/vice-president-communications.tex}

\subsection{Authority and Responsibility of the Executive}
\label{authority-and-responsibility-of-the-executive}
\subsubsection{Executive Authority}
\label{executive-authority}

The Executive has the following authority:

\begin{enumerate}
 \item
  To administer the day-to-day business and operations of the MES.
 \item
  To act as an advisory board to the MES Council on the effects of the MES Council's policy decisions on the day-to-day operations of the MES.
 \item
  To perform the functions of the MES Council during the summer, subject to any restrictions placed on it by the MES Council.
 \item
  In case of an emergency situation where the MES Council is unable to meet, the Executive may make an interim decision, although any policy decisions shall be subject to ratification at the next MES Council meeting.
 \item
  Approve expenditures not exceeding \$500, unless the expenditure is coming from the Executive Operations account.

\end{enumerate}

\subsubsection{Executive Responsibility}
\label{executive-responsibility}
The Executive shall have the following responsibility:

\begin{enumerate}
 \item
  To review the MES budget prepared by the VPF for final recommendations to the MES Council.
 \item
  To form ad-hoc committees when required for the purpose of appointing members to assist in administering the policies of the MES and undertaking new initiatives.
 \item
  To be accountable and transparent for all its actions to the MES Council and to the general membership at each General Meeting.
 \item
  Initiate and encourage new events and initiatives of the MES.
 \item
  Supervise appointed MES Team positions to provide advice or assistance.
 \item
  Liaise with other members or groups within the university community.
 \item
  Promote professionalism within the MES Team.
 \item
  Take a proactive role in the long term goal setting of the MES.
 \item
  Ensure effective transition and continuity between the outgoing Team and the incoming Team, through documentation and orientation.
 \item
  Create a positive and effective working environment.
 \item
  Executive members to hold at least 2 hours of scheduled office hours weekly.

\end{enumerate}

\subsection{Directors}
\label{directors}

\subsubsection{Director of Events}
\label{director-of-events}

The Director of Events shall:

\begin{enumerate}
 \item
  Oversee the organization of social events of the MES foremost, but not limited to:

  \begin{enumerate}
   \item
    Eng Pub Nights
  \end{enumerate}
 \item
  Initiate at least one new activity and ensure the overall schedule of events offers a variety of activities that will appeal to many different groups.
 \item
  Coordinate with the VPC to promote MES activities.
 \item
  Provide the Administrator with dates and details of all events and initiatives.
 \item
  Report to Vice President, Student Life.
  \end{enumerate}
\end{enumerate}

\subsubsection{Director of Clubs}
\label{director-of-clubs}
The Director of Clubs shall:

\begin{enumerate}
 \item
  Meet with MES Groups to support them for Affiliation, Ratification or for other reasons as needed. 
 \item    
  Provide MES Groups bookings for Hatch Junction and Atrium, JHE, and ABB Tower space. 
 \item    
  Collaborate with University Health and Safety and ensure all MES Groups abide by its guidelines. 
 \item
  Promote MES flagship by encouraging the use of MES branding amongst MES groups.
 \item
  Chair the Club Leaders Committee (see MES Bylaws Section \ref{club-leaders-committee}).
 \item
  Serve on the Hatch Operations Committee (Faculty Committee). 
 \item
  Organize Clubsfest to facilitate connections between students and clubs
 \item    
  Organize and update the MES Groups list and bylaws as needed. 
 \item    
  Report to Vice President, Internal.

\end{enumerate}

\subsubsection{Director of Academic Resources}
\label{director-of-academic-resources}
The Director of Academic Resources shall:

\begin{enumerate}
 \item
  Organize all academically related events run by the MES including, but not limited to:

  \begin{enumerate}
   \item
    MES test preparation sessions
   \item
    Technical skills sessions
  \end{enumerate}
 \item
  Be responsible for maintaining all aspects of the McMaster Engineering Society Textbook Library, foremost:

  \begin{enumerate}
   \item
    Actively working to procure new textbooks
   \item
    Maintaining an updated inventory of all textbooks in the MES' possession.
   \item
    Coordinating with the Drain Coordinators to ensure proper instruction is given to the Drain volunteers on how to run the Textbook Library.
   \item
    Cleaning the Textbook Library and ensuring it is kept organized.
   \item
    Ensuring all textbooks in the Textbook Library are labelled with the MES logo.
   \item
    Ensuring all textbooks are returned to the Textbook Library by the end of each semester.
  \end{enumerate}
 \item
  Work with the VPA to plan Back2School Week
 \item
  Chair the Academic Services Committee (See MES Bylaws Section \ref{academic-services-committee})
 \item
  Actively seek ways to improve the academically related events and services provided by the MES.
 \item
  Organize and maintain the Notes/Test Bank.
 \item
  Report to Vice President, Academic.
\end{enumerate}

\subsubsection{Director of Operations}
\label{director-of-operations}
The Director of Operations shall: 

\begin{enumerate} 
  \item   
   Supervise the care and usage of all MES Equipment Rental (see MES Bylaws Section 11.7). 
  \item    
   Be responsible for ensuring that MES Groups using the equipment rentals are aware of how to safely handle them. 
  \item    
   Ensure all MES Groups renting equipment rentals complete the appropriate forms and provide the appropriate deposits. 
  \item    
   Assist VPI and GHC Coordinator(s) in organizing space allocations for MES Clubs and Teams in the Gerald Hatch Centre. 
  \item     
   Maintain the MES Google Workspace and the email service and mailing lists. 
  \item    
   Take appropriate action to protect passwords for the MES mailing lists and any other MES passwords. 
  \item    
   Report to VPI. 
  \item    
   Assist in supervision of the following positions: 
  
  \being{enumerate} 
   \item    
    DW Lounge Coordinator(s) 
   \item     
    Gerald Hatch Centre Student Coordinator(s)    
   \item    
    Trailer Coordinator(s)   
  \end{enumerate} 
  
\subsection{Representatives}
\label{representatives}
\subsubsection{Program
 Representatives}
\label{program-representatives}

\begin{enumerate}
 \item
  There shall be one representative from each program listed below, who is a member of that program.

  \begin{enumerate}
   \item
    Chemical Engineering
   \item
    Civil Engineering
   \item
    Software Engineering
   \item
    Electrical and Computer Engineering
   \item
    Engineering Physics
   \item
    Materials Science and Engineering
   \item
    Mechanical Engineering
   \item
    Mechatronics Engineering
   \item
    Computer Science
   \item
    Engineering and Management
   \item
    Engineering and Society
   \item
    Bachelor of Technology
   \item
    Integrated Biomedical Engineering and Health Sciences
  \end{enumerate}

 \item
  Responsibilities:

  \begin{enumerate}
   \item
    To act as a liaison between the MES, their respective program societies, and the students represented by those societies.
   \item
    To keep the MES informed of all program and program society issues and activities.
   \item
    Advertise all MES events and activities to their respective constituents by making class announcements, distributing posters from the Graphic Designers, distributing MES publications, and promoting the MES website.
   \item
    Actively encourage participation in all MES activities.
   \item
    Ensure quorum is met at General Meetings
   \item
    Serve as a representative for their program on the macLAB Board of Directors (see MES Bylaws Section \ref{maclab-board-of-directors}).
   \item
    Organize their program's team(s) for MES Sports Tournaments, with the exception of Management and Society.
   \item
    Represent the MES at their respective Program Society's meetings.
   \item
    Report to the President.
  \end{enumerate}

\end{enumerate}

\subsubsection{First Year
 Representatives}
\label{first-year-representatives}
\begin{enumerate}
 \item
  Four Level I Representatives, who are full members of the MES, will be elected from the first year class that they represent, one from each first year program:

  \begin{enumerate}
   \item
    Engineering 1
   \item
    Computer Science
   \item
    Bachelor of Technology
   \item
    iBioMed
  \end{enumerate}
 \item
  Responsibilities:

  \begin{enumerate}
   \item
    Advertise all MES events and activities to first year students by making class announcements, promoting the MES website.
   \item
    Serve on the First Year Council (see MES Bylaws Section \ref{first-year-committee}).
   \item
    \begin{enumerate}
     \item
      The Engineering 1 Representative also serves and is a voting member on the MES Council.
    \end{enumerate}
   \item
    Represent the concerns and opinions of first year students within the Faculty of Engineering.
   \item
    Promote and encourage first year attendance at all events.
   \item
    Promote MES committees and activities to the first year students.
   \item
    Work with the VPA to raise concerns of first year students with appropriate course coordinators.
   \item
    Assist in the recruitment and transition of subsequent First Year Representatives.
   \item
    Organize first year student team(s) for the Ratboy Soccer Tournament.
   \item
    Report to the Director of First Year Council.
   \item
    If applicable, represent the MES and First Year Council at their respective program society's meetings.
  \end{enumerate}
\end{enumerate}

\subsubsection{Executive Transition Process}
\label{executive-transition-process}
\begin{enumerate}
 \item
  The outgoing and incoming Executive shall plan and attend a retreat to make long term plans for the year ahead and carry out a more detailed transition.
 \item
  The President is to organize, plan and document the retreat.
 \item
  The budget for this retreat shall be set for no more than \$2000, taken from the Executive Operations budget line.
 \item
  The incoming Executive shall prepare a presentation for the first Council meeting following the retreat, outlining what was accomplished on the retreat, what the budget was used for, and informing Council of the year's objectives and long-term goals.
 \item
  The Outgoing Executive shall remain engaged with the Incoming Executive over the summer months to impart institutional knowledge, share best practices, and facilitate an effective transition.
 \item
  The Outgoing Vice-President Finance shall oversee the year-end closing and ensure the completion of the audit for their fiscal year under their purview.
 \item
  The Outgoing President and Outgoing Vice-President External Relations shall attend ESSCO PM.
\end{enumerate}

