\section{Special Events}
\label{special-events}
All MES events must follow the principles and procedures set out by the
University Health and Safety (UHS).

\subsection{Welcome Week}
\label{welcome-week}

WW is a week-long event for first year students aimed at welcoming them into the McMaster Engineering community. WW is filled with programs and events relating to community building, wellness, and McMaster Engineering traditions \& culture. %% Welcome Week (hereafter ``WW")

``Welcome Week Representatives'' will hereafter be referred to as
``Welcome Week Reps''.

\subsubsection{Purpose}
\label{purpose-1}

\begin{enumerate}
 \item
  WW has the following purposes:

  \begin{enumerate}
   \item
    To welcome first year engineering students to McMaster.
   \item
    To foster pride in our school and our faculty.
   \item
    To introduce first years to their new classmates and to upper year students.

  \end{enumerate}
\end{enumerate}

\subsubsection{Welcome Week Rules}
\label{welcome-week-rules}
\begin{enumerate}
 \item
  The purpose of Welcome Week Reps is to make sure first year students have a fun and safe WW.
 \item
  Reps shall remain sober while participating in official WW activities.
 \item
  Reps are responsible for the safety of first year students during all WW events.
 \item
  Reps are responsible for ensuring that first year students return safely to their place of residence.
 \item
  Reps must adhere to the Welcome Week Discipline Code, as well as the principles of the University's Risk Management Policy.
 \item
  Reps will ensure that they demonstrate respect for first year students at all times.
 \item
  Any Reps displaying offensive and/or inappropriate behaviour may be removed from their position, without warning, upon agreement by the Welcome Week Orientation Planning Committee. This includes any violation of their Welcome Week contract.
\end{enumerate}

\subsubsection{Organization}
\label{organization}
WW is to be planned out by the appointed Co-OCs and Welcome Week
Orientation Planning Committee. The Coordinators and Committee select
the Welcome Week Reps for that year.

\begin{enumerate}
 \item
  See MES Bylaws Section \ref{co-orientation-coordinators} for responsibilities of the Co-OCs.
 \item
  See MES Bylaws Section \ref{welcome-week-orientation-planning-committee} for responsibilities of the Welcome Week Orientation Planning Committee.
 \item
  Selection of Welcome Week Reps:

  \begin{enumerate}
   \item
    The Welcome Week Orientation Planning Committee will hold group auditions and individual interviews for all interested engineering students in February or early March.
   \item
    Common sense and safety questions are to be asked individually of the group members at the time of their try-out. All chosen Welcome Week Reps must display a commitment to abiding by the rules of Student Success Center and the Welcome Week Values and Guiding Principles.
   \item
    Upon selection, Welcome Week Reps must hold an MES Membership for the academic year for which selection occurs.
   \item
    A financial need exemption to the MES Membership may be granted at the discretion of the Co-OCs.
  \end{enumerate}
 \item
  Welcome Week Rep Suits:

  \begin{enumerate}
   \item
    Welcome Week Rep suits are red coveralls, with the McMaster Engineering Fireball on the back.
   \item
    Welcome Week Reps must pay for their own suits.
   \item
    A financial need exemption to the cost of the Welcome Week Rep suit may be granted at the discretion of the Co-OCs.
   \item
    Decoration of the suits must be in good taste, and comply with the guidelines of the Student Success Centre.

  \end{enumerate}
\end{enumerate}

\subsubsection{Events}
\label{events}
\begin{enumerate}
 \item
  All Welcome Week Reps and first year students must respect the community and environment they are in, especially during off campus events.
 \item
  Events that may be held during WW include:

  \begin{enumerate}
   \item
    Success Portfolio
   \item
    Engineering Olympics
   \item
    Faculty Night
   \item
    Bus Pull

    \begin{enumerate}
     \item
      A double-length city bus is pulled around downtown to raise money for Cystic Fibrosis as part of Shinerama.
    \end{enumerate}
   \item
    Campus Tour
   \item
    Trust Walk
   \item
    Graduation Ceremony

    \begin{enumerate}
     \item
      First years are knighted by the MES President with the Sword, and congratulated by the Welcome Week Reps.
    \end{enumerate}
   \item
    WW video
  \end{enumerate}
\end{enumerate}

\subsection{Fireball}
\label{fireball}
\begin{enumerate}
 \item
  Fireball is to be organized by the Fireball Coordinator(s) and scheduled to occur between the second last week of January and the first week of February. The Fireball venue should be booked by the Fireball Coordinator(s) shortly after they are ratified into the position.
 \item
  Tables will be reserved for the Faculty, upon request, at the same cost as the student tickets.
 \item
  The outgoing MES President shall present the President's Sword to the incoming MES President at the event.
 \item
  MES Council Appreciation Awards, ``Image of an Engineer'' Awards, President's Awards, Outstanding Teaching Assistant Awards, and Faculty Appreciation Awards shall be presented at the event (see MES Policy Manual Section C). % more

\end{enumerate}

\subsection{Back2School Week}
\label{back2school-week}
\begin{enumerate}
 \item
  Back2School Week is held two weeks following WW.
 \item
  It is the VPSL's duty to organize the event with the assistance of the VPA, DoE, DoAR, and DoC.
 \item
  Back2School Week will include the ClubsFest, Fresh Pub, and Hatch Open House.
\end{enumerate}

\subsection{Other Events}
\label{other-events}

\subsubsection{Eng Pub Nights}
\label{eng-pub-nights}
\begin{enumerate}
 \item
  Eng Pub Nights are organized by the DoE (see MES Bylaws Section \ref{director-of-events})
  List of tentative Eng Pub Nights, along with the month which they are planned for should be decided by the end of June. These dates are to be given to Handbook Editor(s) for inclusion in the Engineering Handbook for the following year.
\end{enumerate}
